DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-09-26 2015CompleteDerivation/completeDerivation.tex
Feedback from Christine

\begin{rationale}
	The idea is that the system internal dynamics has no effect on the motion and therefore can be disregarded. Consider the motion of our cannonball under gravitational and inertial forces: its temperature or precise chemical composition does not affect the overall trajectory. We can assume kinematic equivalence. Consider the motion of a helicopter under gravitational and inertial forces: the pilot and the amount of fuel in the tank do affect the trajectory. We cannot assume kinematic equivalence.

	Kinematic equivalence can't therefore be assumed for all systems. Yet, the assumption is fundamental because the particles of any material will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not hold.

	The first natural question to ask is: why does state evolution always describe a trajectory in space? It's because spatial extent is the one way in which all materials can be decomposed. As we divide materials in smaller pieces, physical location is one state variable we can assign to the particles of all materials. If we assume kinematic equivalence, the kinematic variables will be enough to identify a particle state. In such a case, we will call the material \emph{minimal} to highlight the idea that the state is described by the smallest possible number of d.o.f..

	When the system is not minimal, the state of the particles will be described by some additional independent state variables. The overall dynamics will be the spatial d.o.f. coupled with the internal d.o.f.. Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials. In other words: any material is \emph{at least} a minimal material and will possess the dynamic needed to describe its motion in space.

@@

We first note that the system of units for the trajectories is fully defined by $\mathcal{M}$. Any trajectory can be specified once a set of coordinates $x^\alpha$ is chosen. Under the kinematic assumption, these units must also be sufficient to describe the dynamics. Therefore $\mathcal{M}$ has conceptually two roles: it is the space where the motion unfolds and it is the space that defines the system of units for our states. This already tells us that the our state space is a hyper-surface of $T^*\mathcal{M}$.

@@

Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $q^\alpha = (t, q^i) = x^\alpha$. This already tells us that the trajectory is differentiable since $q^i(t)$ is differentiable. The four velocity $u^\alpha = d_s x^\alpha$ is therefore well defined and it can be used to identify trajectories. The magnitude of the four velocity can be rescaled by changing the evolution parameter, $u^\alpha = d_s x^\alpha = d_s \hat{s} \, d_{\hat{s}} x^\alpha = d_s \hat{s} \, \hat{u}^\alpha$, and is therefore not physical. Only the direction is physical so the four velocity gives us another $n$ independent variables to identify trajectories. The combined position and velocity, the kinematic variables, identify $2n + 1$ trajectories using a hyper-surface of the tangent bundle $T\mathcal{M}$. This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$. We cannot add more curves, as a homeomorphism preserves dimensionality, therefore we identified the suitable space for the trajectories of particles of a classical material under kinematic equivalence.

@@

	We claim there exists a homeomorphism between $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ and a hyper-surface of $T\mathcal{M}$. Let $\mathcal{s} \in \bar{\mathcal{S}}$ be the particle state for a minimal classical material. Let $x^\alpha(s)$ the evolution associated with that state. Let $t(\mathcal{s})$ be the instant of time at which $\mathcal{s}$ is defined. $x^\alpha(s(t(\mathcal{s}))) = f(\mathcal{s}) = q^\alpha$ as the initial position cannot be derived by the law of evolution. $q^\alpha$ is differentiable therefore $d_s x^\alpha$ exists. Suppose $d_s x^\alpha$ is not part of the state. $d_s x^\alpha = d_s q^\alpha$ is determined by the law of evolution. $x^\alpha(s)$ is determined by the law of evolution. $q^\alpha$ determines the trajectory $x^\alpha(s)$ and the state $\mathcal{s}$, as the material is minimal. This is a contradiction: $\mathcal{s}$ is not identified by just $q^\alpha$. $u^\alpha=d_s x^\alpha$ is part of the state. $p_\alpha = p_\alpha(x^\alpha, u^\alpha)$. The map $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is continuous by \ref{continuous_map} and is bijective as the material is minimal. It is a homemorphism between $T^*\mathcal{M}$ and $T\mathcal{M}$. $\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$ therefore its image is a hyper-surface of $T\mathcal{M}$.

>>>>

\begin{rationale}
	The idea is that the system internal dynamics has no effect on the overall motion and therefore can be disregarded. Consider the motion of our cannonball under gravitational and inertial forces: its temperature or precise chemical composition does not affect the overall trajectory. We can assume kinematic equivalence. Suppose we now add magnetic forces to our consideration. The magnetization of the cannonball may now affect the motion and therefore we cannot always reconstruct the full state by simply looking at the trajectory. We cannot assume kinematic equivalence.

	Kinematic equivalence can't therefore be assumed for all systems. Yet, the assumption is fundamental because the particles of any material, during their evolution, will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not hold.

	The first natural question to ask is: why does state evolution of a particle of a classical material always describe a trajectory in space? It's because spatial extent is the one way in which all materials can be decomposed. As we divide materials into smaller pieces, physical location is one state variable we can assign to the particles of all materials. If we assume kinematic equivalence, the kinematic variables will be enough to identify a particle state. In such a case, we will call the material \emph{minimal} to highlight the idea that the state is described by the smallest possible number of d.o.f..

	When the system is not minimal, the state of the particles will be described by some additional independent state variables. The overall dynamics will be the spatial d.o.f. coupled with the internal d.o.f.. Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials. In other words: any material is an extension to a minimal material.

@@

We first note that the system of units for the trajectories is fully defined by $\mathcal{M}$. Any trajectory can be specified once a set of coordinates $x^\alpha$ is chosen. Under the kinematic assumption, these units must also be sufficient to describe the dynamics. Therefore $\mathcal{M}$ has conceptually two roles: it is the space where the motion unfolds and it is the space that defines the system of units for our states. This already tells us that the state space for the particles of the material is a hyper-surface of $T^*\mathcal{M}$.

@@

Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $q^\alpha = (t, q^i) = x^\alpha$. This already tells us that the trajectory is differentiable since $q^\alpha(s)$ is differentiable. The four velocity $u^\alpha = d_s x^\alpha$ is therefore well defined and it can be used to identify trajectories. The magnitude of the four velocity can be rescaled by changing the evolution parameter, $u^\alpha = d_s x^\alpha = d_s \hat{s} \, d_{\hat{s}} x^\alpha = d_s \hat{s} \, \hat{u}^\alpha$, and is therefore not physical. Only the direction is physical so the four velocity gives us another $n$ independent variables to identify trajectories. The combined position and velocity, the kinematic variables, identify $2n + 1$ trajectories using a hyper-surface of the tangent bundle $T\mathcal{M}$. This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$. We cannot add more curves, as a homeomorphism preserves dimensionality, therefore we identified the suitable space for the trajectories of particles of a classical material under kinematic equivalence.

@@

	We claim there exists a homeomorphism between $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ and a hyper-surface of $T\mathcal{M}$. Let $\mathcal{s} \in \bar{\mathcal{S}}$ be the particle state for a minimal classical material. Let $x^\alpha(s)$ be the evolution associated with that state. Let $t(\mathcal{s})$ be the instant of time at which $\mathcal{s}$ is defined. $q^\alpha(\mathcal{s}) \equiv x^\alpha(s(t(\mathcal{s})))$ are state variables as they are real function of the state. $q^\alpha$ is differentiable therefore $d_s x^\alpha$ exists. Suppose $d_s x^\alpha$ is not part of the state. $d_s x^\alpha = d_s q^\alpha$ is determined by the law of evolution. $x^\alpha(s)$ is determined by the law of evolution. $q^\alpha$ determines the trajectory $x^\alpha(s)$ and the state $\mathcal{s}$, as the material is minimal. This is a contradiction: $\mathcal{s}$ is not identified by just $q^\alpha$. $u^\alpha=d_s x^\alpha$ is part of the state. $p_\alpha = p_\alpha(x^\alpha, u^\alpha)$. The map $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is continuous by \ref{continuous_map} and is bijective as the material is minimal. It is a homemorphism between $T^*\mathcal{M}$ and $T\mathcal{M}$. $\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$ therefore its image is a hyper-surface of $T\mathcal{M}$.
